\mytitle{}

### 2019-10-01 paper/paper-revised.tex
Final paper and in blue!

\subsection{Contributions}
\cbstart
This paper contributes two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
These are designed to make effective decisions irrespective of protocol, and act on individual flows at the edge of any network topology.
We offer an in-depth investigation into suitable features for automatic DDoS mitigation, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
These features have been suggested by past studies, and independently tested in their own contexts.
Our study is the first attempt to quantify the individual efficacy of each in an RL setting.

We implement reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
To our knowledge, this is the first attempt to study or replicate Opus-based VoIP traffic, which has become commonplace since the codec's release in 2012.
These new traffic models inform an empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
We additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS \cbend defences.

@@

\subsection{System Design and Assumptions}
A deployment environment is a network with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
These nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model/design, and control the proportion of upstream packets from each external host to discard.
Each destination node $s$ has a maximum capacity, \cbend $U_s$.

@@

\cbstart
Our system design applies to both software-defined and traditional networks of arbitrary shape and size.
Only the ingress/egress nodes from a network need to be OpenFlow-enabled, as it is advantageous to perform filtering as close to a flow's source as possible.
In a traditional network, each agent has exclusive control over its switch's tables.
In a fully software-defined network, these agents require exclusive control over the first table, forwarding legitimate packets to subsequent tables managed by the network's controller.
The main difference is that a traditional network needs this additional hardware, and does not allow an operator to dynamically determine where the "edge" of their network lies through vNF relocation.
\cbend

@@

%?? How is global state acquired? See \cref{fig:global-state-path}. Why take paths in the way that we do? Mention deterministic ECMP routing etc...
Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}) depending upon the bandwidth measurements regularly received from monitors in the environment.
For any flow, an agent then computes the path it would take through the network.
The incoming load recorded along the first hop, last hop, and tertiles of the path may then be tile-coded together.
\cbstart In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) the load measurement of a middle hop or the last hop. \cbend
\Cref{fig:global-state-path} illustrates the process.

@@

We describe and analyse each of the per-flow features included in the state vector throughout \cref{sec:rethinking-the-state-space}.
Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
\cbstart Rather than having the network push the data to an agent, the agent requests this information about active flows periodically to isolate it from non-control-plane traffic and to eliminate the risk of resource exhaustion by excessive requests. \cbend

@@

Each agent begins with an empty work list.
For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
Uncompleted work is passed on to the next timestep, until the list is emptied, at which point it is repopulated using the set of available measurements.
To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields.
Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 3 flows in this time.
We expect this should be tuned based on the frequency at which statistics arrive.
\cbstart
Naturally, this implies that an agent must carry work forward (and coalesce state updates) when the \emph{host density} $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.
\cbend

@@

\subsection{TCP (HTTP) Traffic Model}\label{sec:tcp-http-traffic-model}
%?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
To model legitimate TCP traffic, server nodes run an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
Hosts and clients both use TCP Cubic \cite{rfc8312}.
Each host's download rate is limited to match whatever maximum bandwidth was assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.).
On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime.
This presents a more balanced distribution of flow duration and size, with large files included to model elephant flows.

@@

%?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
\subsection{Attack Traffic Model}\label{sec:attack-traffic-model}
\cbstart
Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
We configure each instance of \emph{hping3} to generate ethernet MTU-sized packets (\SI{1500}{\byte}) with a random source and destination port towards a target server, and configure the output rate $r$ (in \si{\mega\bit\per\second}) by setting the inter-arrival time $t_{\mathit{attack}}=\frac{1500 \cdot 8}{r\cdot10^6}$.
This fulfils certain characteristics of many types of amplification DDoS traffic: it is congestion-unaware \cite{DBLP:conf/ndss/Rossow14}, and packets are larger than the minimum frame size and identically-sized (e.g., NTP amplification traffic is fragmented at the application layer into \SI{482}{\byte} chunks \cite{cisco-ntp-amp}).
We differ from NTP amplification in frame size so that inter-arrival times are larger, to keep emulation of the network feasible at high traffic rates.
\cbend

@@

We compare our work most naturally against MARL, introduced by \textcite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
We are most interested in seeing how their approach contrasts with ours across different topologies and workloads.
Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network---we aim to see how these characteristics affect performance and learning rate.
\cbstart
Marl is known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} strategy, yet the state of the art has long since moved on.
To paint a more current picture, we compare our work against an effective modern approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
SPIFFY tests a proportion of flows by routing them through an alternate path with higher bandwidth, observing how their speed changes some time later.
This comparison lets us position our new agent designs against the state of the art, observing that SPIFFY has a similar mode of interaction to RL-based systems (taking action, observing an effect, and acting once again) and does not rely on protocol characteristics or signatures.
We make the simplifying assumption that a suitable unused path exists (with identical bandwidth to the server's link).
We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\! \times$} expansion from legitimate flows, making a judgement after \cbend \SI{5}{\second}.

To test this, we made use of both traffic models introduced in \cref{sec:a-new-normal} (OPUS and TCP), both topologies discussed below (1-dest vs Fat-Tree), and vary the amount of hosts typically communicating over each agent's ingress/egress node.
Additionally, we evaluated our new models in multi-agent mode (\emph{separate}, no model sharing), and in single-agent mode (\emph{single}, 0-cost perfect information sharing).
In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between episodes---a host is malicious with probability $\operatorname{P}(\mathit{malicious})$, and is benign otherwise.
Benign hosts generate traffic from either \cref{sec:tcp-http-traffic-model,sec:udp-opus-traffic-model} depending on the experiment, while malicious hosts generate traffic according to \cref{sec:attack-traffic-model} (both at experiment-dependent rates).

@@

Benign hosts each occupied \SIrange{0}{1}{\mega\bit\per\second}, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
%The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the absence of a suitable heuristic classifier $g(\cdot)$, disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and high densities of hosts at egress points.
%?? Why? ...
%Of these, the latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
%These are important issues, particularly when we consider real-world deployment.
%Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
%?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE PROTOCOLS
Malicious hosts each sent \SIrange{2.5}{6}{\mega\bit\per\second} when attacking UDP traffic, though this was increased to \SIrange{4}{7}{\mega\bit\per\second} when using TCP-like traffic (to meaningfully impact benign flows).
Given $n$ and $\operatorname{P}(\mathit{malicious})$, we see an expected malicious bandwidth \numrange{1.27}{1.87} and \SIrange{2.03}{2.18}{$\! \times U_s$} respectively.
%The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
\cbstart For our choices of $n$ in both topologies, we observe $N_{\mathit{hosts}} \in \left\{24, 48, 96, 192\right\}$, and an expected number of malicious hosts $\mathbb{E}\left[N_{\mathit{attackers}}\right] \in \left\{9.6, 19.2, 38.4, 76.8\right\}$.
For the largest choice of $n$, we see an expected total attack traffic $\mathbb{E}\left[V_{\mathit{attack}}\right] = 334.05$ and \SI{422.4}{\mega\bit\per\second} for Opus and HTTP traffic respectively.
\cbend

@@

We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) and \emph{SPIFFY} under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
%?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
%Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---providing a rough idea of expected performance, with the highest-performing model in bold and the best RL-based model underlined.
Average rewards take into account any portions of time that an agent allows illegal system states.
Several plots augment this, illustrating peak performance or the amount of time which an agent requires to learn.

@@

In a single-destination network, we observe that Marl's performance degrades as $n$ increases.
Typically, our \emph{Instant} agent design achieves the best performance in multi-agent mode, having lower collateral damage than the current state-of-the-art, but sharply degrades at low $n$ when agents share experience.
This trend reverses for the \emph{Guarded} model, which improves as $n$ increases and in single-agent mode---when $n\ge4$, the single-agent variant offers consistent improvement.
%Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
\Cref{fig:udp-tree-box} shows the preserved traffic in multi-agent mode.
%?? Discuss multi-dest topol once all results available.
When defending multiple destinations, we see a sharp decrease in the effectiveness of all agent designs.
Our new agent designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
\cbstart Interestingly, SPIFFY is unable to effectively protect constant bitrate traffic. \cbend

@@

\Cref{tab:av-vals} shows that Marl offers a low (though fairly consistent) level of protection for TCP traffic, which the \emph{Instant} agent offers no substantial improvement over.
However, \emph{Guarded} agents offer a remarkable improvement for this class of traffic, particularly when experience can be shared---offering a \SI{2.21}{$\!\times$} improvement over the state-of-the art during training, which is made clearer in \cref{fig:tcp-tree-box}.
\Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$\!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
%SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
%?? Discuss multi-dest topol once results available.
We observe that the same trends present themselves in the multi-destination topology: \emph{Guarded} remains the best fit for TCP, in both training modes.
Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
In both cases, SPIFFY greatly outperforms the RL-based methods.

@@

\subsection{Increased Attack Volume}\label{sec:results-attack-volume}
\cbstart
To assess the effect of larger volumes of attack traffic, we increase an attacker's output by various factors, supposing $n=16$ with HTTP traffic (\emph{Guarded}, Single); \cref{tab:atk-vol} records the expected rate of attack and average performance.
The initial increase in traffic volume causes the steepest reduction in performance (due to the increased cost of incorrect action), though performance levels out as attack traffic \cbend increases.

\begin{table}
\cbstart
	\centering
%	\vspace{-0.1cm}
	\caption{Average reward versus attack volume.\label{tab:atk-vol}}
\resizebox{0.45\linewidth}{!}{
\begin{tabular}{@{}SSS@{}}
	\toprule\multicolumn{1}{c}{Factor} & \multicolumn{1}{c}{$\mathbb{E}\left[V_{\mathit{attack}}\right]$ (\si{\mega\bit\per\second})} & \multicolumn{1}{c}{Reward} \\ \midrule
	2.0 & 844.8 & 0.625 \\
	2.5 & 1056.0 & 0.620 \\
	3.0 & 1267.2 & 0.619 \\
	3.5 & 1478.4 & 0.600 \\
	\bottomrule
\end{tabular}
}
\cbend
\vspace{-0.75cm}
\end{table}

@@

%Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
Measurements from each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second} per flow.
This is reassuring when measured alongside the insights from other work.
\Textcite{DBLP:conf/sigcomm/ChenL0L18} observe that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
%Most flows are short, and flow-size follows a heavy-tailed distribution.
That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.
\cbstart
This cost is constant and independent of network size.
As discussed in \cref{sec:systems-considerations}, we are able to judge 3 flows before this deadline: the difference is primarily accounted for by serialisation/communication delays and single-threaded processing in the Python language.
\cbend
%?? TODO: update with modern numbers...

@@

%Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
\cbstart
Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this strand still trails behind existing (exact) DDoS flow detection mechanisms where TCP traffic is concerned.
The ability to better protect VoIP traffic when compared against one of these approaches is a curious observation, which suggests that other (exact) protocol-agnostic approaches may carry hidden assumptions and is a promising direction for future investigation.
Similar traffic makes up a significant fraction of network load today (\SIrange{18}{27}{\percent}).
Although we have conducted work to map the territory, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
The benefits we have at present are, however, substantial.
What we offer above many of the approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments, low-overhead and fixed-cost decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
Moreover, our decision making processes are entirely agnostic of the protocol or content of traffic, offering future-proofing against the introduction of new transports.
\cbend

>>>>

\subsection{Contributions}
\cbstart
{\color{uofgcobalt}
This paper contributes two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
These are designed to make effective decisions irrespective of protocol, and act on individual flows at the edge of any network topology.
We offer an in-depth investigation into suitable features for automatic DDoS mitigation, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
These features have been suggested by past studies, and independently tested in their own contexts.
Our study is the first attempt to quantify the individual efficacy of each in an RL setting.

We implement reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
To our knowledge, this is the first attempt to study or replicate Opus-based VoIP traffic, which has become commonplace since the codec's release in 2012.
These new traffic models inform an empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
We additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS \cbend defences.}

@@

\subsection{System Design and Assumptions}
A deployment environment is a network with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
These nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
{\color{uofgcobalt}\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model/design}, and control the proportion of upstream packets from each external host to discard.
Each destination node $s$ has a maximum capacity, \cbend $U_s$.

@@

\cbstart
{\color{uofgcobalt}
Our system design applies to both software-defined and traditional networks of arbitrary shape and size.
Only the ingress/egress nodes from a network need to be OpenFlow-enabled, as it is advantageous to perform filtering as close to a flow's source as possible.
In a traditional network, each agent has exclusive control over its switch's tables.
In a fully software-defined network, these agents require exclusive control over the first table, forwarding legitimate packets to subsequent tables managed by the network's controller.
The main difference is that a traditional network needs this additional hardware, and does not allow an operator to dynamically determine where the ``edge'' of their network lies through vNF relocation.}
\cbend

@@

%?? How is global state acquired? See \cref{fig:global-state-path}. Why take paths in the way that we do? Mention deterministic ECMP routing etc...
Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}) depending upon the bandwidth measurements regularly received from monitors in the environment.
For any flow, an agent then computes the path it would take through the network.
The incoming load recorded along the first hop, last hop, and tertiles of the path may then be tile-coded together.
\cbstart {\color{uofgcobalt}In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) the load measurement of a middle hop or the last hop.} \cbend
\Cref{fig:global-state-path} illustrates the process.

@@

We describe and analyse each of the per-flow features included in the state vector throughout \cref{sec:rethinking-the-state-space}.
Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
\cbstart {\color{uofgcobalt}Rather than having the network push the data to an agent, the agent requests this information about active flows periodically to isolate it from non-control-plane traffic and to eliminate the risk of resource exhaustion by excessive requests.} \cbend

@@

Each agent begins with an empty work list.
For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
Uncompleted work is passed on to the next timestep, until the list is emptied, at which point it is repopulated using the set of available measurements.
To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields.
Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 3 flows in this time.
We expect this should be tuned based on the frequency at which statistics arrive.
\cbstart
{\color{uofgcobalt}Naturally, this implies that an agent must carry work forward (and coalesce state updates) when \emph{host density} is $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.}
\cbend

@@

\subsection{TCP (HTTP) Traffic Model}\label{sec:tcp-http-traffic-model}
%?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
To model legitimate TCP traffic, server nodes run an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
Hosts and clients both use TCP Cubic \cite{rfc8312}.
Each host's download rate is limited to match the maximum bandwidth assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.) \cbstart {\color{uofgcobalt}as a browser \cbend might.}
On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime.
This presents a balanced distribution of flow duration and size, with large files included to model elephant flows.

@@

%?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
\subsection{Attack Traffic Model}\label{sec:attack-traffic-model}
\cbstart
{\color{uofgcobalt}Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
We configure each instance of \emph{hping3} to generate ethernet MTU-sized packets (\SI{1500}{\byte}) with a random source and destination port towards a target server, and configure the output rate $r$ (in \si{\mega\bit\per\second}) by setting the inter-arrival time $t_{\mathit{attack}}=\frac{1500 \cdot 8}{r\cdot10^6}$.
This fulfils certain characteristics of many types of amplification DDoS traffic: it is congestion-unaware \cite{DBLP:conf/ndss/Rossow14}, and packets are larger than the minimum frame size and identically-sized (e.g., NTP amplification traffic is fragmented at the application layer into \SI{482}{\byte} chunks \cite{cisco-ntp-amp}).
We differ from NTP amplification in frame size so that inter-arrival times are larger, to keep emulation of the network feasible at high traffic rates.}
\cbend

@@

We compare our work most naturally against MARL, introduced by \textcite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
We are most interested in seeing how their approach contrasts with ours across different topologies and workloads.
Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network---we aim to see how these characteristics affect performance and learning rate.
\cbstart
{\color{uofgcobalt}Marl is known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} strategy, yet the state of the art has long since moved on.
To paint a more current picture, we compare our work against an effective modern approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
SPIFFY tests a proportion of flows by routing them through an alternate path with higher bandwidth, observing how their speed changes some time later.
This comparison lets us position our new agent designs against the state of the art, observing that SPIFFY has a similar mode of interaction to RL-based systems (taking action, observing an effect, and acting once again) and does not rely on protocol characteristics or signatures.
We make the simplifying assumption that a suitable unused path exists (with identical bandwidth to the server's link).
We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\! \times$} expansion from legitimate flows, making a judgement after \cbend \SI{5}{\second}.}

To test this, we made use of both traffic models introduced in \cref{sec:a-new-normal} (OPUS and TCP), both topologies discussed below (1-dest vs Fat-Tree), and vary the amount of hosts typically communicating over each agent's ingress/egress node.
Additionally, we evaluated our new models in multi-agent mode (\emph{separate}, no model sharing), and in single-agent mode (\emph{single}, 0-cost perfect information sharing).
In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between \cbstart episodes---{\color{uofgcobalt}a host is malicious with probability $\operatorname{P}(\mathit{malicious})$, and is benign \cbend otherwise.}
Benign hosts generate traffic from either \cref{sec:tcp-http-traffic-model,sec:udp-opus-traffic-model} depending on the experiment, while malicious hosts generate traffic according to \cref{sec:attack-traffic-model} (both at experiment-dependent rates).

@@

Benign hosts each occupied \SIrange{0}{1}{\mega\bit\per\second}, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
%The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the absence of a suitable heuristic classifier $g(\cdot)$, disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and high densities of hosts at egress points.
%?? Why? ...
%Of these, the latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
%These are important issues, particularly when we consider real-world deployment.
%Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
%?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE PROTOCOLS
Malicious hosts each sent \SIrange{2.5}{6}{\mega\bit\per\second} when attacking UDP traffic, though this was increased to \SIrange{4}{7}{\mega\bit\per\second} when using TCP-like traffic (to meaningfully impact benign flows).
Given $n$ and $\operatorname{P}(\mathit{malicious})$, we see an expected malicious bandwidth \numrange{1.27}{1.87} and \SIrange{2.03}{2.18}{$\! \times U_s$} respectively.
%The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
{\color{uofgcobalt}\cbstart For our choices of $n$ in both topologies, we observe $N_{\mathit{hosts}} \in \left\{24, 48, 96, 192\right\}$, and an expected number of malicious hosts $\mathbb{E}\left[N_{\mathit{attackers}}\right] \in \left\{9.6, 19.2, 38.4, 76.8\right\}$.
For the largest choice of $n$, we see an expected total attack traffic $\mathbb{E}\left[V_{\mathit{attack}}\right] = 334.05$ and \SI{422.4}{\mega\bit\per\second} for Opus and HTTP traffic respectively.
\cbend}

@@

We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) and {\color{uofgcobalt} \cbstart \emph{SPIFFY} under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
%?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
%Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---providing a rough idea of expected performance, with the highest-performing model in bold and the best RL-based model \cbend underlined.}
Average rewards take into account any portions of time that an agent allows illegal system states.
Several plots augment this, illustrating peak performance or the amount of time which an agent requires to learn.

@@

In a single-destination network, we observe that Marl's performance degrades as $n$ increases.
Typically, our \emph{Instant} agent design achieves the best performance in multi-agent mode, having lower collateral damage than the current state-of-the-art, but sharply degrades at low $n$ when agents share experience.
This trend reverses for the \emph{Guarded} model, which improves as $n$ increases and in single-agent mode---when $n\ge4$, the single-agent variant offers consistent improvement.
%Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
\Cref{fig:udp-tree-box} shows the preserved traffic in multi-agent mode.
%?? Discuss multi-dest topol once all results available.
When defending multiple destinations, we see a sharp decrease in the effectiveness of all agent designs.
Our new agent designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
\cbstart {\color{uofgcobalt}Interestingly, SPIFFY is unable to effectively protect constant bitrate traffic.} \cbend

@@

\Cref{tab:av-vals} shows that Marl offers a low (though fairly consistent) level of protection for TCP traffic, which the \emph{Instant} agent offers no substantial improvement over.
However, \emph{Guarded} agents offer a remarkable improvement for this class of traffic, particularly when experience can be shared---offering a \SI{2.21}{$\!\times$} improvement over the state-of-the art during training, which is made clearer in \cref{fig:tcp-tree-box}.
\Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$\!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
%SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
%?? Discuss multi-dest topol once results available.
We observe that the same trends present themselves in the multi-destination topology: \emph{Guarded} remains the best fit for TCP, in both training modes.
Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
{\color{uofgcobalt}In both cases, SPIFFY greatly outperforms the RL-based methods.}

@@

\subsection{Increased Attack Volume}\label{sec:results-attack-volume}
\cbstart
{\color{uofgcobalt}To assess the effect of larger volumes of attack traffic, we increase an attacker's output by various factors, supposing $n=16$ with HTTP traffic (\emph{Guarded}, Single); \cref{tab:atk-vol} records the expected rate of attack and average performance.
The initial increase in traffic volume causes the steepest reduction in performance (due to the increased cost of incorrect action), though performance levels out as attack traffic \cbend increases.}

\begin{table}
\cbstart
	\centering
%	\vspace{-0.1cm}
	\caption{Average reward versus attack volume.\label{tab:atk-vol}}
\resizebox{0.45\linewidth}{!}{
\begin{tabular}{@{}SSS@{}}
	\toprule\multicolumn{1}{c}{Factor} & \multicolumn{1}{c}{$\mathbb{E}\left[V_{\mathit{attack}}\right]$ (\si{\mega\bit\per\second})} & \multicolumn{1}{c}{Reward} \\ \midrule
	1.5 & 633.6 & 0.671 \\
	2.0 & 844.8 & 0.625 \\
	2.5 & 1056.0 & 0.620 \\
	3.0 & 1267.2 & 0.619 \\
	3.5 & 1478.4 & 0.600 \\
	\bottomrule
\end{tabular}
}
\cbend
\vspace{-0.75cm}
\end{table}

@@

%Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
Measurements from each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second} per flow.
This is reassuring when measured alongside the insights from other work.
\Textcite{DBLP:conf/sigcomm/ChenL0L18} observe that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
%Most flows are short, and flow-size follows a heavy-tailed distribution.
That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.
\cbstart
{\color{uofgcobalt}This cost is constant and independent of network size.
As discussed in \cref{sec:systems-considerations}, we are able to judge 3 flows before this deadline: the difference is primarily accounted for by serialisation/communication delays and single-threaded processing in the Python language.}
\cbend
%?? TODO: update with modern numbers...

@@

%Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
\cbstart
{\color{uofgcobalt}Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this strand still trails behind existing (exact) DDoS flow detection mechanisms where TCP traffic is concerned.
The ability to better protect VoIP traffic when compared against one of these approaches is a curious observation, which suggests that other (exact) protocol-agnostic approaches may carry hidden assumptions and is a promising direction for future investigation.
Similar traffic makes up a significant fraction of network load today (\SIrange{18}{27}{\percent}).
Although we have conducted work to map the territory, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
The benefits we have at present are, however, substantial.
What we offer above many of the approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments, low-overhead and fixed-cost decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
Moreover, our decision making processes are entirely agnostic of the protocol or content of traffic, offering future-proofing against the introduction of new transports.}
\cbend
